---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load file:
```{r}
adult<-read.csv(file= "https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep = ",", header=FALSE)

colnames(adult) <- c("Age","workclass","fnlwgt","education","education-num","marital-status","occupation","relationship","race","sex","capital-gain","capital-loss", "hours-per-week", "native-country", "class")

str(adult)
summary(adult)
```
##Age Variable
```{r}
# Histogram
hist(adult$Age, freq = TRUE, col = "white", main = "Age")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$Age, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
`````






```{r}
sd(adult$Age)
var(adult$Age)

```
#Education Num Variable
```{r}
# Histogram
hist(adult$`education-num`, freq = TRUE, col = "white", main = "Education-Num")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$Age, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}

sd(adult$`education-num`)
var(adult$`education-num`)
```
##Capital Gain Variable
```{r}
# Histogram
hist(adult$'capital-gain', freq = TRUE, col = "white", main = "Capital Gain")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$'capital-gain', horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}
sd(adult$`capital-gain`)
var(adult$`capital-gain`)
```

#Capital Loss Variable
```{r}
# Histogram
hist(adult$'capital-loss', freq = TRUE, col = "white", main = "Capital Loss")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$'capital-loss', horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```


```{r}
sd(adult$`capital-loss`)
var(adult$`capital-loss`)

```
#Hours per week Variable
```{r}
# Histogram
hist(adult$`hours-per-week`, freq = TRUE, col = "white", main = "Hours per Week")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$`hours-per-week`, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}
sd(adult$`hours-per-week`)
var(adult$`hours-per-week`)

```

#Workclass Variable
```{r}
table(adult$workclass)

countsworkclass <- table(adult$workclass)
barplot(countsworkclass, main="Workclass Distribution",
        xlab="Workclass")
```

#Education Variable
```{r}
table(adult$education)

countsedu <- table(adult$education)
barplot(countsedu, main="Education Distribution",
        xlab="Education Level")
```

#Marital Status Variable
```{r}
table(adult$`marital-status`)

countsmarital <- table(adult$`marital-status`)
barplot(countsmarital, main="Marital Status Distribution",
        xlab="Marital Status")
```
#Occupation Variable
```{r}
table(adult$occupation)

countsoccu <- table(adult$occupation)
barplot(countsoccu, main="Occupation Distribution",
        xlab="Occupation")

```

#Relationship Variable
```{r}
table(adult$relationship)

countsrel <- table(adult$relationship)
barplot(countsrel, main="Relationship Distribution",
        xlab="Relationship")
```

#Race Variable
```{r}
table(adult$race)

countsrace <- table(adult$race)
barplot(countsrace, main="Race Distribution",
        xlab="Race")
```
#Sex Variable
```{r}
table(adult$sex)

countssex <- table(adult$sex)
barplot(countssex, main="Sex Distribution",
        xlab="Sex")

```
#Native Country Variable
```{r}
table(adult$`native-country`)

countsnative <- table(adult$`native-country`)
barplot(countsnative, main="Native Country Distribution",
        xlab="Country")

```
#Create duplicate file for cleaning
```{r}
adultclean<-adult

str(adultclean)
```

##FIND MISSING VALUES
```{r}
# find elements
replace <- adultclean == " ?"
# replace elements with NA
is.na(adultclean) <- replace
```


```{r}
#Missing values per attribute
sapply(adultclean, function(x) sum(is.na(x)))

#subset of missing values
new_DF <- adultclean[rowSums(is.na(adultclean)) > 0,]
```
##Clean white space
```{r}
adultclean <- as.data.frame(apply(adultclean,2, function(x) gsub("\\s+", "", x))) 
```


#Fill Missing Values with Mode
```{r}
#Replace Workclass missing values with mode - Private
A<-is.na(adultclean$workclass)
adultclean$workclass[A]<-"Private"

#Replace occupation missing values with mode - Prof Specialty
B<-is.na(adultclean$occupation)
adultclean$occupation[B]<-"Prof-specialty"


#Replace Native Country missing values with mode - United-States
C<-is.na(adultclean$`native-country`)
adultclean$`native-country`[C]<-"United-States"
```

# Redo counts after missing values
```{r}
#Workclass Variable
table(adultclean$workclass)

#Occupation Variable
table(adultclean$occupation)

#Native Country  Variable
table(adultclean$`native-country`)
```

#Look at Class Variable
```{r}
table(adult$class)
```

###Compare Education and Education-num
```{r}
library(dplyr)
distinct(adult,education,`education-num`)



```
#Drop Education-num
```{r}
adultclean$`education-num`<-NULL
```

#Drop Fnlwgt 
```{r}
adultclean$fnlwgt<-NULL
```

#Update Workclass Categories
```{r}
adultclean["workclass"][adultclean["workclass"]=='Federal-gov' |adultclean["workclass"]=='Local-gov' |adultclean["workclass"]=='State-gov']<-"Government"

adultclean["workclass"][adultclean["workclass"]=='Never-worked' |adultclean["workclass"]=='Without-pay']<-"Other"

adultclean["workclass"][adultclean["workclass"]=='Self-emp-inc' |adultclean["workclass"]=='Self-emp-not-inc']<-"Self Employed"

table(adultclean$workclass)


countswc <- table(adultclean$workclas)
barplot(countswc, main="Workclass Distribution",
        xlab="Workclass")
```
#Update Education Categories
```{r}
adultclean["education"][adultclean["education"]=='Preschool' |adultclean["education"]=='1st-4th' |adultclean["education"]=='5th-6th' |adultclean["education"]=='7th-8th' |adultclean["education"]=='9th' |adultclean["education"]=='10th' |adultclean["education"]=='11th' |adultclean["education"]=='12th']<-"Below Highschool"

adultclean["education"][adultclean["education"]=='HS-grad']<-"Highschool Grad"

adultclean["education"][adultclean["education"]=='Assoc-acdm' |adultclean["education"]=='Assoc-voc']<-"Associate"

adultclean["education"][adultclean["education"]=='Prof-school' |adultclean["education"]=='Doctorate' |adultclean["education"]=='Masters']<-"Grad School"

table(adultclean$education)
countsed <- table(adultclean$education)
barplot(countsed, main="Education Distribution",
        xlab="Education")
```
#Clean Marital Status
```{r}
adultclean["marital-status"][adultclean["marital-status"]=='Married-civ-spouse' |adultclean["marital-status"]=='Married-spouse-absent' |adultclean["marital-status"]=='Married-AF-spouse']<-"Married"

table(adultclean$`marital-status`)
```
#Clean Native Country
```{r}
adultclean["native-country"][adultclean["native-country"]=='Cuba' |adultclean["native-country"]=='Jamaica' |adultclean["native-country"]=='India' |adultclean["native-country"]=='Mexico' |adultclean["native-country"]=='South' |adultclean["native-country"]=='Puerto-Rico' |adultclean["native-country"]=='Honduras' |adultclean["native-country"]=='England'|adultclean["native-country"]=='Canada'|adultclean["native-country"]=='Germany'|adultclean["native-country"]=='Iran'|adultclean["native-country"]=='Philippines'|adultclean["native-country"]=='Italy'|adultclean["native-country"]=='Poland'|adultclean["native-country"]=='Columbia'|adultclean["native-country"]=='Cambodia'|adultclean["native-country"]=='Thailand'|adultclean["native-country"]=='Ecuador'|adultclean["native-country"]=='Laos'|adultclean["native-country"]=='Taiwan'|adultclean["native-country"]=='Haiti'|adultclean["native-country"]=='Portugal'|adultclean["native-country"]=='Dominican-Republic'|adultclean["native-country"]=='El-Salvador'|adultclean["native-country"]=='France'|adultclean["native-country"]=='Guatemala'|adultclean["native-country"]=='China'|adultclean["native-country"]=='Japan'|adultclean["native-country"]=='Yugoslavia'|adultclean["native-country"]=='Peru'|adultclean["native-country"]=='Outlying-US(Guam-USVI-etc)'|adultclean["native-country"]=='Scotland'|adultclean["native-country"]=='Trinadad&Tobago'|adultclean["native-country"]=='Greece'|adultclean["native-country"]=='Nicaragua'|adultclean["native-country"]=='Vietnam'|adultclean["native-country"]=='Hong'|adultclean["native-country"]=='Ireland'|adultclean["native-country"]=='Hungary'|adultclean["native-country"]=='Holand-Netherlands']<-"Outside US"
```


#Correlation

```{r}

```

```{r}
subset_adult<-adultclean[,c("Age","hours-per-week","capital-gain","capital-loss")]
plot(subset_adult)

subset_adult$Age<-as.integer(subset_adult$Age)
subset_adult$`hours-per-week`<-as.integer(subset_adult$`hours-per-week`)
subset_adult$`capital-gain`<-as.integer(subset_adult$`capital-gain`)
subset_adult$`capital-loss`<-as.integer(subset_adult$`capital-loss`)
str(subset_adult)

cor(subset_adult,method="pearson")
cor(subset_adult,method="spearman")
```


```{r}
subset_adult$Age<-as.integer(subset_adult$Age)
subset_adult$`hours-per-week`<-as.integer(subset_adult$`hours-per-week`)
subset_adult$`capital-gain`<-as.integer(subset_adult$`capital-gain`)
subset_adult$`capital-loss`<-as.integer(subset_adult$`capital-loss`)
str(subset_adult)
```


```{r}

adultclean$Age<-as.integer(adultclean$Age)
adultclean$`hours-per-week`<-as.integer(adultclean$`hours-per-week`)
adultclean$`capital-gain`<-as.integer(adultclean$`capital-gain`)
adultclean$`capital-loss`<-as.integer(adultclean$`capital-loss`)
str(adultclean)
```

#Update Class Variable
#<=50K is 0 and >50K is 1
```{r}


adultclean["class"][adultclean["class"]=='<=50K']<-0
adultclean["class"][adultclean["class"]=='>50K']<-1

```


##Check Outliers for Hours per Week



```{r}
adultclean$`hours-per-week`<-as.integer(adultclean$`hours-per-week`)
adultclean$zscorehours<-scale(adultclean$`hours-per-week`)
length(which(abs(adultclean$zscorehours)>3))
```

#Replace outliers with mean value
```{r}
adultclean$`hours-per-week`[which(abs(adultclean$zscorehours)>3)] <- mean(adultclean$`hours-per-week`, na.rm = TRUE)


# Histogram
hist(adultclean$`hours-per-week`, freq = TRUE, col = "white", main = "Hours per Week")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adultclean$`hours-per-week`, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```
#Standardize Numeric Variables
```{r}
normalize <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}

adultnorm<-adultclean

adultnorm$`capital-gain`<-as.integer(adultnorm$`capital-gain`)
adultnorm$`capital-loss`<-as.integer(adultnorm$`capital-loss`)
adultnorm$Age<-as.integer(adultnorm$Age)
adultnorm$class<-as.integer(adultnorm$class)
#str(adultnorm)
adultnorm$`capital-gain`<-normalize(adultnorm$`capital-gain`)
adultnorm$`capital-loss`<-normalize(adultnorm$`capital-loss`)
adultnorm$Age<-normalize(adultnorm$Age)
adultnorm$`hours-per-week`<-normalize(adultnorm$`hours-per-week`)

adultnorm$zscorehours<-NULL
```

#Split into Test and Train
```{r}
set.seed(13)
train_index=sample(1:nrow(adultnorm),0.7*nrow(adultnorm))
train<-adultnorm[train_index,]
test<-adultnorm[-train_index,]

head(train)
head(test)
```

#Logistic Regression
```{r}
glm_model<-glm(class~.,train,family="binomial")
summary(glm_model)
```
#Logistic Regression Confusion Matrix
```{r}
lrpredicted <- predict(glm_model,test,type="response")
predicted_income<-ifelse(lrpredicted>=0.5,1,0)
confusionmatrix<-table(actual=test$class,predicted=predicted_income)
print(confusionmatrix)

Accuracy<-sum(diag(confusionmatrix))/nrow(test)
print(Accuracy)
```
#####Dummy Variables

```{r}
install.packages('fastDummies')
library('fastDummies')
```


```{r}
adultdum<-adultnorm
adultdum <- dummy_cols(adultdum, select_columns = 'workclass')
adultdum <- dummy_cols(adultdum, select_columns = 'education')
adultdum <- dummy_cols(adultdum, select_columns = 'marital-status')
adultdum <- dummy_cols(adultdum, select_columns = 'occupation')
adultdum <- dummy_cols(adultdum, select_columns = 'relationship')
adultdum <- dummy_cols(adultdum, select_columns = 'race')
adultdum <- dummy_cols(adultdum, select_columns = 'sex')

adultdum <- dummy_cols(adultdum, select_columns = 'native-country')

adultdum$workclass<-NULL
adultdum$education<-NULL
adultdum$`marital-status`<-NULL
adultdum$occupation<-NULL
adultdum$relationship<-NULL
adultdum$race<-NULL
adultdum$sex<-NULL
adultdum$`native-country`<-NULL
```


```{r}
#Convert all to numeric
adultdum2 <- data.frame(lapply(adultdum, function(x) as.numeric(as.character(x))))

```


##KNN
```{r}
install.packages("class")
install.packages("gmodels")
library(class)
library(gmodels)
```


```{r}
set.seed(13)
train_indexdum=sample(1:nrow(adultdum2),0.7*nrow(adultdum2))
traindum<-adultdum2[train_indexdum,]
testdum<-adultdum2[-train_indexdum,]

train.set.knn<-traindum[-5]
test.set.knn<-testdum[-5]



train.set_labels<-traindum$class
test.set_labels<-testdum$class



test_pred <- knn(train = train.set.knn, test = test.set.knn, cl = train.set_labels, k=10)
test_pred2<-knn(train = train.set.knn, test = test.set.knn, cl = train.set_labels, k=180)
```

```{r}

#k=10
CrossTable(x=test.set_labels, y=test_pred, prop.chisq=FALSE)
table(Actual=test.set_labels, Predicted=test_pred)
```
```{r}
#k=180
CrossTable(x=test.set_labels, y=test_pred2, prop.chisq=FALSE)
table(Actual=test.set_labels, Predicted=test_pred2)
```

##Decision Tree
```{r setup, include=FALSE}
library("party")
```


```{r}
```


```{r}
adult_ctree <- ctree(class ~ Age + capital.gain + capital.loss + hours.per.week + workclass_Government + workclass_Other + workclass_Private + workclass_Self.Employed + education_Associate + education_Bachelors + education_Below.Highschool + education_Grad.School + education_Highschool.Grad + education_Some.college + marital.status_Divorced + marital.status_Married.AF.spouse +marital.status_Married.civ.spouse + marital.status_Married.spouse.absent + marital.status_Never.married + marital.status_Separated + marital.status_Widowed + occupation_Adm.clerical + occupation_Armed.Forces +occupation_Craft.repair + occupation_Exec.managerial + occupation_Farming.fishing + occupation_Handlers.cleaners + occupation_Machine.op.inspct + occupation_Other.service + occupation_Priv.house.serv + occupation_Prof.specialty + occupation_Protective.serv + occupation_Sales + occupation_Tech.support + occupation_Transport.moving + relationship_Husband + relationship_Not.in.family + relationship_Other.relative + relationship_Own.child + relationship_Unmarried + relationship_Wife + race_Amer.Indian.Eskimo + race_Asian.Pac.Islander + race_Black +race_Other + race_White + sex_Female + sex_Male, data=adultdum2)
```

```{r}
print(adult_ctree)
```
```{r}
plot(adult_ctree, type="simple")
```
```{r}
traindec<-traindum
testdec<-testdum

traindec["class"][traindec["class"]==0]<-'<=50K'
traindec["class"][traindec["class"]==1]<-'>50K'

testdec["class"][testdec["class"]==0]<-'<=50K'
testdec["class"][testdec["class"]==1]<-'>50K'

testdec$class<-as.factor(testdec$class)
traindec$class<-as.factor(traindec$class)
```


#Model Decision Tree
```{r}
adult_ctree_model <- ctree(class ~ ., data=traindec)
```

```{r}
adult_ctree_prediction <- predict(adult_ctree_model, newdata=testdec) 
# gives the probability for each class
head(adult_ctree_prediction)
```

```{r}
table(adult_ctree_prediction, testdec$class)
```

###Naive Bayes
```{r}
nb_model<-naiveBayes(class~.,data=train)
```

```{r}
nb_predict<-predict(nb_model,newdata=test)
```


```{r}


table(nb_predict,test$class)
```

