---
title: "Untitled"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load file:
```{r}
adult<-read.csv(file= "https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep = ",", header=FALSE)

colnames(adult) <- c("Age","workclass","fnlwgt","education","education-num","marital-status","occupation","relationship","race","sex","capital-gain","capital-loss", "hours-per-week", "native-country", "class")

str(adult)
summary(adult)
```
##Age Variable
```{r}
# Histogram
hist(adult$Age, freq = TRUE, col = "white", main = "Age")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$Age, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
`````

```{r}
ageclass<-subset(adult,select=c("Age","class"))


ageclass$Age <- cut(
  ageclass$Age,
  breaks = c(0, 30, 45, 60, Inf),
  labels = c("Under 30", "31-45", "46-60", "Over 60"),
  right  = FALSE
)

```



```{r}
ageclasstab<-table(ageclass$class,ageclass$Age)
barplot(ageclasstab, main="Age Distribution",
        xlab="Age",legend.text = rownames(ageclasstab),beside=TRUE)

print(ageclasstab)
```



```{r}
sd(adult$Age)
var(adult$Age)

```
#Education Num Variable
```{r}
# Histogram
hist(adult$`education-num`, freq = TRUE, col = "white", main = "Education-Num")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$Age, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}

sd(adult$`education-num`)
var(adult$`education-num`)
```
##Capital Gain Variable
```{r}
# Histogram
hist(adult$'capital-gain', freq = TRUE, col = "white", main = "Capital Gain")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$'capital-gain', horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}
sd(adult$`capital-gain`)
var(adult$`capital-gain`)
```

#Capital Loss Variable
```{r}
# Histogram
hist(adult$'capital-loss', freq = TRUE, col = "white", main = "Capital Loss")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$'capital-loss', horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```


```{r}
sd(adult$`capital-loss`)
var(adult$`capital-loss`)

```
#Hours per week Variable
```{r}
# Histogram
hist(adult$`hours-per-week`, freq = TRUE, col = "white", main = "Hours per Week")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adult$`hours-per-week`, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```

```{r}
sd(adult$`hours-per-week`)
var(adult$`hours-per-week`)

```

#Workclass Variable
```{r}
table(adult$workclass)

countsworkclass <- table(adult$workclass)
barplot(countsworkclass, main="Workclass Distribution",
        xlab="Workclass")
```

#Education Variable
```{r}
table(adult$education)

countsedu <- table(adult$education)
barplot(countsedu, main="Education Distribution",
        xlab="Education Level")
```

#Marital Status Variable
```{r}
table(adult$`marital-status`)

countsmarital <- table(adult$`marital-status`)
barplot(countsmarital, main="Marital Status Distribution",
        xlab="Marital Status")
```
#Occupation Variable
```{r}
table(adult$occupation)

countsoccu <- table(adult$occupation)
barplot(countsoccu, main="Occupation Distribution",
        xlab="Occupation")

```

#Relationship Variable
```{r}
table(adult$relationship)

countsrel <- table(adult$relationship)
barplot(countsrel, main="Relationship Distribution",
        xlab="Relationship")
```

#Race Variable
```{r}
table(adult$race)

countsrace <- table(adult$race)
barplot(countsrace, main="Race Distribution",
        xlab="Race")

raceclass<-table(adult$class,adult$race)
barplot(raceclass, main="Race Distribution",
        xlab="Race",legend.text = rownames(raceclass),beside=TRUE)

print(raceclass)
```
#Sex Variable
```{r}
table(adult$sex)

countssex <- table(adult$sex)
barplot(countssex, main="Sex Distribution",
        xlab="Sex")

sexclass<-table(adult$class,adult$sex)
barplot(sexclass, main="Sex Distribution",
        xlab="Sex",legend.text = rownames(sexclass),beside=TRUE)

print(sexclass)


```
#Native Country Variable
```{r}
table(adult$`native-country`)

countsnative <- table(adult$`native-country`)
barplot(countsnative, main="Native Country Distribution",
        xlab="Country")

```
#Create duplicate file for cleaning
```{r}
adultclean<-adult

str(adultclean)
```

##FIND MISSING VALUES
```{r}
# find elements
replace <- adultclean == " ?"
# replace elements with NA
is.na(adultclean) <- replace
```


```{r}
#Missing values per attribute
sapply(adultclean, function(x) sum(is.na(x)))

#subset of missing values
new_DF <- adultclean[rowSums(is.na(adultclean)) > 0,]
```
##Clean white space
```{r}
adultclean <- as.data.frame(apply(adultclean,2, function(x) gsub("\\s+", "", x))) 
```


#Fill Missing Values with Mode
```{r}
#Replace Workclass missing values with mode - Private
A<-is.na(adultclean$workclass)
adultclean$workclass[A]<-"Private"

#Replace occupation missing values with mode - Prof Specialty
B<-is.na(adultclean$occupation)
adultclean$occupation[B]<-"Prof-specialty"


#Replace Native Country missing values with mode - United-States
C<-is.na(adultclean$`native-country`)
adultclean$`native-country`[C]<-"United-States"
```

# Redo counts after missing values
```{r}
#Workclass Variable
table(adultclean$workclass)

#Occupation Variable
table(adultclean$occupation)

#Native Country  Variable
table(adultclean$`native-country`)
```

#Look at Class Variable
```{r}
table(adult$class)
```

###Compare Education and Education-num
```{r}
library(dplyr)
distinct(adult,education,`education-num`)



```
#Drop Education-num
```{r}
adultclean$`education-num`<-NULL
```

#Drop Fnlwgt 
```{r}
adultclean$fnlwgt<-NULL
```

#Update Workclass Categories
```{r}
adultclean["workclass"][adultclean["workclass"]=='Federal-gov' |adultclean["workclass"]=='Local-gov' |adultclean["workclass"]=='State-gov']<-"Government"

adultclean["workclass"][adultclean["workclass"]=='Never-worked' |adultclean["workclass"]=='Without-pay']<-"Other"

adultclean["workclass"][adultclean["workclass"]=='Self-emp-inc' |adultclean["workclass"]=='Self-emp-not-inc']<-"Self Employed"

table(adultclean$workclass)


countswc <- table(adultclean$workclas)
barplot(countswc, main="Workclass Distribution",
        xlab="Workclass")
```
#Update Education Categories
```{r}
adultclean["education"][adultclean["education"]=='Preschool' |adultclean["education"]=='1st-4th' |adultclean["education"]=='5th-6th' |adultclean["education"]=='7th-8th' |adultclean["education"]=='9th' |adultclean["education"]=='10th' |adultclean["education"]=='11th' |adultclean["education"]=='12th']<-"Below Highschool"

adultclean["education"][adultclean["education"]=='HS-grad']<-"Highschool Grad"

adultclean["education"][adultclean["education"]=='Assoc-acdm' |adultclean["education"]=='Assoc-voc']<-"Associate"

adultclean["education"][adultclean["education"]=='Prof-school' |adultclean["education"]=='Doctorate' |adultclean["education"]=='Masters']<-"Grad School"

table(adultclean$education)
countsed <- table(adultclean$education)
barplot(countsed, main="Education Distribution",
        xlab="Education")
```
#Clean Marital Status
```{r}
adultclean["marital-status"][adultclean["marital-status"]=='Married-civ-spouse' |adultclean["marital-status"]=='Married-spouse-absent' |adultclean["marital-status"]=='Married-AF-spouse']<-"Married"

table(adultclean$`marital-status`)
```
#Clean Native Country
```{r}
adultclean["native-country"][adultclean["native-country"]=='Cuba' |adultclean["native-country"]=='Jamaica' |adultclean["native-country"]=='India' |adultclean["native-country"]=='Mexico' |adultclean["native-country"]=='South' |adultclean["native-country"]=='Puerto-Rico' |adultclean["native-country"]=='Honduras' |adultclean["native-country"]=='England'|adultclean["native-country"]=='Canada'|adultclean["native-country"]=='Germany'|adultclean["native-country"]=='Iran'|adultclean["native-country"]=='Philippines'|adultclean["native-country"]=='Italy'|adultclean["native-country"]=='Poland'|adultclean["native-country"]=='Columbia'|adultclean["native-country"]=='Cambodia'|adultclean["native-country"]=='Thailand'|adultclean["native-country"]=='Ecuador'|adultclean["native-country"]=='Laos'|adultclean["native-country"]=='Taiwan'|adultclean["native-country"]=='Haiti'|adultclean["native-country"]=='Portugal'|adultclean["native-country"]=='Dominican-Republic'|adultclean["native-country"]=='El-Salvador'|adultclean["native-country"]=='France'|adultclean["native-country"]=='Guatemala'|adultclean["native-country"]=='China'|adultclean["native-country"]=='Japan'|adultclean["native-country"]=='Yugoslavia'|adultclean["native-country"]=='Peru'|adultclean["native-country"]=='Outlying-US(Guam-USVI-etc)'|adultclean["native-country"]=='Scotland'|adultclean["native-country"]=='Trinadad&Tobago'|adultclean["native-country"]=='Greece'|adultclean["native-country"]=='Nicaragua'|adultclean["native-country"]=='Vietnam'|adultclean["native-country"]=='Hong'|adultclean["native-country"]=='Ireland'|adultclean["native-country"]=='Hungary'|adultclean["native-country"]=='Holand-Netherlands']<-"Outside US"
```


#Correlation

```{r}

```

```{r}
subset_adult<-adultclean[,c("Age","hours-per-week","capital-gain","capital-loss")]
plot(subset_adult)

subset_adult$Age<-as.integer(subset_adult$Age)
subset_adult$`hours-per-week`<-as.integer(subset_adult$`hours-per-week`)
subset_adult$`capital-gain`<-as.integer(subset_adult$`capital-gain`)
subset_adult$`capital-loss`<-as.integer(subset_adult$`capital-loss`)
str(subset_adult)

cor(subset_adult,method="pearson")
cor(subset_adult,method="spearman")
```

```{r}
install.packages("reshape2")
library(reshape2)
 
# creating correlation matrix
corr_mat <- round(cor(subset_adult),2)
 
# reduce the size of correlation matrix
melted_corr_mat <- melt(corr_mat)
head(melted_corr_mat)
 
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var2, Var1, label = value),
          color = "black", size = 4)

```


```{r}
subset_adult$Age<-as.integer(subset_adult$Age)
subset_adult$`hours-per-week`<-as.integer(subset_adult$`hours-per-week`)
subset_adult$`capital-gain`<-as.integer(subset_adult$`capital-gain`)
subset_adult$`capital-loss`<-as.integer(subset_adult$`capital-loss`)
str(subset_adult)
```


```{r}

adultclean$Age<-as.integer(adultclean$Age)
adultclean$`hours-per-week`<-as.integer(adultclean$`hours-per-week`)
adultclean$`capital-gain`<-as.integer(adultclean$`capital-gain`)
adultclean$`capital-loss`<-as.integer(adultclean$`capital-loss`)
str(adultclean)
```

#Update Class Variable
#<=50K is 0 and >50K is 1
```{r}


adultclean["class"][adultclean["class"]=='<=50K']<-0
adultclean["class"][adultclean["class"]=='>50K']<-1

```


##Check Outliers for Hours per Week



```{r}
adultclean$`hours-per-week`<-as.integer(adultclean$`hours-per-week`)
adultclean$zscorehours<-scale(adultclean$`hours-per-week`)
length(which(abs(adultclean$zscorehours)>3))
```

#Replace outliers with mean value
```{r}
adultclean$`hours-per-week`[which(abs(adultclean$zscorehours)>3)] <- mean(adultclean$`hours-per-week`, na.rm = TRUE)


# Histogram
hist(adultclean$`hours-per-week`, freq = TRUE, col = "white", main = "Hours per Week")
# Add new plot
par(new = TRUE)
# Box plot
boxplot(adultclean$`hours-per-week`, horizontal = TRUE, axes = FALSE, col = rgb(0, 0.8, 1, alpha = 0.5))
# Box around the plots
box()
```
#Standardize Numeric Variables
```{r}
normalize <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}

adultnorm<-adultclean

adultnorm$`capital-gain`<-as.integer(adultnorm$`capital-gain`)
adultnorm$`capital-loss`<-as.integer(adultnorm$`capital-loss`)
adultnorm$Age<-as.integer(adultnorm$Age)
adultnorm$class<-as.integer(adultnorm$class)
#str(adultnorm)
adultnorm$`capital-gain`<-normalize(adultnorm$`capital-gain`)
adultnorm$`capital-loss`<-normalize(adultnorm$`capital-loss`)
adultnorm$Age<-normalize(adultnorm$Age)
adultnorm$`hours-per-week`<-normalize(adultnorm$`hours-per-week`)
adultnorm$zscorehours<-NULL
```



```{r}
colnames(adultnorm)
names(adultnorm)[names(adultnorm) == "marital-status"] <- "maritalstatus"
names(adultnorm)[names(adultnorm) == "capital-gain"] <- "capitalgain"
names(adultnorm)[names(adultnorm) == "capital-loss"] <- "capitalloss"
names(adultnorm)[names(adultnorm) == "hours-per-week"] <- "hoursperweek"
names(adultnorm)[names(adultnorm) == "native-country"] <- "nativecountry"

colnames(adultnorm)
```

#Split into Test and Train
```{r}
set.seed(13)
train_index=sample(1:nrow(adultnorm),0.7*nrow(adultnorm))
train<-adultnorm[train_index,]
test<-adultnorm[-train_index,]

set.seed(14)
train_index2=sample(1:nrow(adultnorm),0.7*nrow(adultnorm))
train2<-adultnorm[train_index2,]
test2<-adultnorm[-train_index2,]

set.seed(15)
train_index3=sample(1:nrow(adultnorm),0.7*nrow(adultnorm))
train3<-adultnorm[train_index3,]
test3<-adultnorm[-train_index3,]




head(train)
head(test)

head(train2)
head(test2)
```

#Logistic Regression
```{r}
start.timelr <- Sys.time()
glm_model<-glm(class~.,train,family="binomial")
summary(glm_model)
end.timelr <- Sys.time()
time.takenlr <- end.timelr - start.timelr
time.takenlr
```
#Logistic Regression 2
```{r}
start.timelr2 <- Sys.time()
glm_model2<-glm(class~.,train2,family="binomial")
summary(glm_model2)
end.timelr2 <- Sys.time()
time.takenlr2 <- end.timelr2 - start.timelr2
time.takenlr2
```
#Logistic Regression 3
```{r}
start.timelr3 <- Sys.time()
glm_model3<-glm(class~.,train3,family="binomial")
summary(glm_model3)
end.timelr3 <- Sys.time()
time.takenlr3 <- end.timelr3 - start.timelr3
time.takenlr3
```

#Logistic Regression Confusion Matrix
```{r}
lrpredicted <- predict(glm_model,test,type="response")
predicted_income<-ifelse(lrpredicted>=0.5,1,0)
confusionmatrix<-table(actual=test$class,predicted=predicted_income)
print(confusionmatrix)

library(caret)
Accuracy<-sum(diag(confusionmatrix))/nrow(test)
print(Accuracy)
sensitivity(confusionmatrix)
specificity(confusionmatrix)
```
#Logistic Regression Confusion Matrix2
```{r}
lrpredicted2 <- predict(glm_model2,test2,type="response")
predicted_income2<-ifelse(lrpredicted2>=0.5,1,0)
confusionmatrix2<-table(actual=test2$class,predicted=predicted_income2)
print(confusionmatrix2)

library(caret)
Accuracy2<-sum(diag(confusionmatrix2))/nrow(test2)
print(Accuracy2)
sensitivity(confusionmatrix2)
specificity(confusionmatrix2)
```
#Logistic Regression Confusion Matrix3
```{r}
lrpredicted3 <- predict(glm_model3,test3,type="response")
predicted_income3<-ifelse(lrpredicted3>=0.5,1,0)
confusionmatrix3<-table(actual=test3$class,predicted=predicted_income3)
print(confusionmatrix3)

library(caret)
Accuracy3<-sum(diag(confusionmatrix3))/nrow(test2)
print(Accuracy3)
sensitivity(confusionmatrix3)
specificity(confusionmatrix3)
```
###Logistic Regression Predict White Male
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="White", sex="Male", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```
###Logistic Regression Predict White Female
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="White", sex="Female", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```
###Logistic Regression Predict Black Male
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="Black", sex="Male", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```

###Logistic Regression Predict Black female
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="Black", sex="Female", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```
###Logistic Regression Predict Asian Male
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="Asian-Pac-Islander", sex="Male", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```


###Logistic Regression Predict Indian Male
```{r}
predict(glm_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="Amer-Indian-Eskimo", sex="Male", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="response")
```
#####Dummy Variables

```{r}
#install.packages('fastDummies')
library('fastDummies')
```


```{r}
adultdum<-adultnorm
adultdum <- dummy_cols(adultdum, select_columns = 'workclass')
adultdum <- dummy_cols(adultdum, select_columns = 'education')
adultdum <- dummy_cols(adultdum, select_columns = 'maritalstatus')
adultdum <- dummy_cols(adultdum, select_columns = 'occupation')
adultdum <- dummy_cols(adultdum, select_columns = 'relationship')
adultdum <- dummy_cols(adultdum, select_columns = 'race')
adultdum <- dummy_cols(adultdum, select_columns = 'sex')

adultdum <- dummy_cols(adultdum, select_columns = 'nativecountry')

adultdum$workclass<-NULL
adultdum$education<-NULL
adultdum$maritalstatus<-NULL
adultdum$occupation<-NULL
adultdum$relationship<-NULL
adultdum$race<-NULL
adultdum$sex<-NULL
adultdum$nativecountry<-NULL
```


```{r}
#Convert all to numeric
adultdum2 <- data.frame(lapply(adultdum, function(x) as.numeric(as.character(x))))

```


##KNN
```{r}
#install.packages("class")
#install.packages("gmodels")
library(class)
library(gmodels)
```


```{r}
set.seed(13)
train_indexdum=sample(1:nrow(adultdum2),0.7*nrow(adultdum2))
traindum<-adultdum2[train_indexdum,]
testdum<-adultdum2[-train_indexdum,]

train.set.knn<-traindum[-5]
test.set.knn<-testdum[-5]



train.set_labels<-traindum$class
test.set_labels<-testdum$class


start.timekn1 <- Sys.time()
test_pred <- knn(train = train.set.knn, test = test.set.knn, cl = train.set_labels, k=10)
end.timekn1 <- Sys.time()
time.takenkn1 <- end.timekn1 - start.timekn1
time.takenkn1

start.timekn2 <- Sys.time()
test_pred2<-knn(train = train.set.knn, test = test.set.knn, cl = train.set_labels, k=180)
end.timekn2 <- Sys.time()
time.takenkn2<- end.timekn2 - start.timekn2
time.takenkn2
```

```{r}

#k=10
CrossTable(x=test.set_labels, y=test_pred, prop.chisq=FALSE)
table(Actual=test.set_labels, Predicted=test_pred)
```
```{r}
#k=180
CrossTable(x=test.set_labels, y=test_pred2, prop.chisq=FALSE)
table(Actual=test.set_labels, Predicted=test_pred2)
```

##KNN2
```{r}
set.seed(19)
train_indexdum2=sample(1:nrow(adultdum2),0.7*nrow(adultdum2))
traindum2<-adultdum2[train_indexdum2,]
testdum2<-adultdum2[-train_indexdum2,]

train.set.knn2<-traindum2[-5]
test.set.knn2<-testdum2[-5]



train.set_labels2<-traindum2$class
test.set_labels2<-testdum2$class



test_pred2a <- knn(train = train.set.knn2, test = test.set.knn2, cl = train.set_labels2, k=10)

test_pred2b<-knn(train = train.set.knn2, test = test.set.knn2, cl = train.set_labels2, k=180)

```

```{r}
#k=10
CrossTable(x=test.set_labels2, y=test_pred2a, prop.chisq=FALSE)
table(Actual=test.set_labels2, Predicted=test_pred2a)
```
```{r}
#k=180
CrossTable(x=test.set_labels2, y=test_pred2b, prop.chisq=FALSE)
table(Actual=test.set_labels2, Predicted=test_pred2b)
```
##KNN3
```{r}
set.seed(29)
train_indexdum3=sample(1:nrow(adultdum2),0.7*nrow(adultdum2))
traindum3<-adultdum2[train_indexdum3,]
testdum3<-adultdum2[-train_indexdum3,]

train.set.knn3<-traindum3[-5]
test.set.knn3<-testdum3[-5]



train.set_labels3<-traindum3$class
test.set_labels3<-testdum3$class



test_pred3a <- knn(train = train.set.knn3, test = test.set.knn3, cl = train.set_labels3, k=10)

test_pred3b<-knn(train = train.set.knn3, test = test.set.knn3, cl = train.set_labels3, k=180)

```

```{r}
#k=10
CrossTable(x=test.set_labels3, y=test_pred3a, prop.chisq=FALSE)
table(Actual=test.set_labels3, Predicted=test_pred3a)
```

```{r}
#k=180
CrossTable(x=test.set_labels3, y=test_pred3b, prop.chisq=FALSE)
table(Actual=test.set_labels3, Predicted=test_pred3b)
```

  

##Decision Tree
```{r }
library("party")
```


```{r}
```


```{r}
start.timedt <- Sys.time()
adult_ctree <- ctree(class ~ ., data=adultdum2)
end.timedt <- Sys.time()
time.takendt <- end.timedt - start.timedt
time.takendt

```

```{r}
start.timedt2 <- Sys.time()
print(adult_ctree)
end.timedt2 <- Sys.time()
time.takendt2 <- end.timedt2 - start.timedt2
time.takendt2
```
```{r}
plot(adult_ctree, type="simple")
```
```{r}
traindec<-traindum
testdec<-testdum

traindec["class"][traindec["class"]==0]<-'<=50K'
traindec["class"][traindec["class"]==1]<-'>50K'

testdec["class"][testdec["class"]==0]<-'<=50K'
testdec["class"][testdec["class"]==1]<-'>50K'

testdec$class<-as.factor(testdec$class)
traindec$class<-as.factor(traindec$class)
```


#Model Decision Tree
```{r}

start.timedt3 <- Sys.time()
adult_ctree_model <- ctree(class ~ ., data=traindec)
end.timedt3 <- Sys.time()
time.takendt3 <- end.timedt3 - start.timedt3
time.takendt3

```

```{r}
adult_ctree_prediction <- predict(adult_ctree_model, newdata=testdec) 
# gives the probability for each class
head(adult_ctree_prediction)
```

```{r}
table(adult_ctree_prediction, testdec$class)
```
#Decision Tree 2

```{r}
traindec2<-traindum2
testdec2<-testdum2

traindec2["class"][traindec2["class"]==0]<-'<=50K'
traindec2["class"][traindec2["class"]==1]<-'>50K'

testdec2["class"][testdec2["class"]==0]<-'<=50K'
testdec2["class"][testdec2["class"]==1]<-'>50K'

testdec2$class<-as.factor(testdec2$class)
traindec2$class<-as.factor(traindec2$class)
```

```{r}
adult_ctree_model2 <- ctree(class ~ ., data=traindec2)
```


```{r}
adult_ctree_prediction2 <- predict(adult_ctree_model2, newdata=testdec2) 
# gives the probability for each class
head(adult_ctree_prediction2)
```
```{r}
table(adult_ctree_prediction2, testdec2$class)
```
#Decision Tree 3

```{r}
traindec3<-traindum3
testdec3<-testdum3

traindec3["class"][traindec3["class"]==0]<-'<=50K'
traindec3["class"][traindec3["class"]==1]<-'>50K'

testdec3["class"][testdec3["class"]==0]<-'<=50K'
testdec3["class"][testdec3["class"]==1]<-'>50K'

testdec3$class<-as.factor(testdec3$class)
traindec3$class<-as.factor(traindec3$class)
```

```{r}
adult_ctree_model3 <- ctree(class ~ ., data=traindec3)
```


```{r}
adult_ctree_prediction3 <- predict(adult_ctree_model3, newdata=testdec3) 
# gives the probability for each class
head(adult_ctree_prediction3)
```
```{r}
table(adult_ctree_prediction3, testdec3$class)
```
###Naive Bayes
```{r}
library(e1071)


start.timenb <- Sys.time()
nb_model<-naiveBayes(class~.,data=train)
end.timenb <- Sys.time()
time.takennb <- end.timenb - start.timenb
time.takennb
```

```{r}
nb_predict<-predict(nb_model,newdata=test)
```


```{r}


table(nb_predict,test$class)
```
##Naive Bayes 2
```{r}

nb_model2<-naiveBayes(class~.,data=train2)

```

```{r}
nb_predict2<-predict(nb_model2,newdata=test2)

```
```{r}
table(nb_predict,test$class)
```

##Naive Bayes 3

```{r}
nb_model3<-naiveBayes(class~.,data=train3)
```

```{r}
nb_predict3<-predict(nb_model3,newdata=test3)
```

```{r}
table(nb_predict3,test3$class)
```

##Naive Bayes Predict
#White Female
```{r}
predict(nb_model, newdata=data.frame(Age=0.5, workclass="Private",education="Highschool Grad", maritalstatus="Married",occupation="Prof-specialty", relationship="Husband",race="Black", sex="Female", capitalgain=0, hoursperweek=0.5, nativecountry="United-States", capitalloss=0), type="raw")
```
```{r}

install.packages('leaps')
library(MASS) # stepwise regression
library(leaps)

subsets<-regsubsets(class~.,
data=adultclean, nbest=1)
sub.sum <- summary(subsets)
as.data.frame(sub.sum$outmat)



```

